# pokemon unite

```{r}
library(scales)
```

## data ingestion

raw data are stored by pokemon level, let's merge them

```{r}
read_stats = function(){
  df = data.frame()
  for(i in 1:15){
    uri = sprintf("datasets/stats/%s.csv", i)
    dfi = read.csv(uri)
    dfi$level = i
    df = rbind(df, dfi)
  }
}
```

```{r}
df = read_stats()
```

```{r}
df_rn = paste(
  toupper(substr(df$name, 1,4)),
  df$level,
  sep="")

row.names(df) = df_rn
df$name = NULL
```

## dim red

let's normalize features a bit

```{r}
for(i in 1:length(df)){
  df[i] = ((df[i] - min(df[i])) /(max(df[i])-min(df[i])))
}

df = df[df$level > (13/15),]
df$level = NULL
```

we cannot use k means over the original point cloud, beacus cause features are not comparable (distances between points are not euclidean, and an appropriate distance has not been found)

k means has been applied to reduce dimensions

```{r}
res = princomp(df, cor=T)
summary(res)
```

```{r}
res$loading
```

```{r}
screeplot(res)
plot(res$scores, cex=0.0)
text(res$scores, rownames(df), cex=0.6)
abline(h=0, v=0)
```

```{r}
pr.var=res$sdev^2
pve=pr.var/sum(pr.var)

plot(cumsum(pve), xlab="Principal Component", ylab="Cumulative Proportion of Variance Explained", ylim=c(0,1),type='b')
```

## clust

```{r}
biplot(res, cex=0.5)
abline(h=0, v=0)
```

```{r}
wssplot <- function(data, nc=15, seed=1234){
  wss = (nrow(data)-1)*sum(apply(data,2,var))
  for (i in 2:nc){
    set.seed(seed)
    wss[i] <- sum(kmeans(data, centers=i)$withinss)}
  plot(1:nc, wss, type="b", xlab="Number of Clusters",
       ylab="Within groups sum of squares")}

wssplot(df[2:length(df)], nc=12) 
```

```{r}
library(cluster)
clusplot(df, kmeans(df, centers=6)$cluster, 
         main='2D representation of the Cluster solution',
         color=TRUE, shade=TRUE,
         labels=2, lines=0, cex=0.5)
```

```{r}
# Compute DBSCAN using fpc package
library("fpc")

db = fpc::dbscan(res$scores, eps = 1.6, MinPts = 2)

# Plot DBSCAN results
library("factoextra")
fviz_cluster(db, data = res$scores, stand = FALSE,
ellipse = TRUE, show.clust.cent = FALSE,
geom = "point",palette = "jco", ggtheme = theme_classic())
```

```{r}
# clookup = data.frame()
# 
# for(c in 1:length(df)){
#   clookup = rbind(clookup, c(c, row.names(df)[c], db$cluster[c]))
# }
# 
# colnames(clookup) = c('id','name', 'cluster')
```

```{r}

```

## prediction

```{r}
dt = read.csv('datasets/matches.csv')
```

### score distro

```{r}
cnt = aggregate(dt$score, by=list(pokemon=dt$pokemon), FUN=length)

cnt[order(-cnt$x),]
```

```{r}
shapiro.test(dt$score)
```

```{r}
library(ggpubr)
ggqqplot(dt$score)
ggdensity(dt, x = "score", fill = "lightgray", title = "score") +
  stat_overlay_normal_density(color = "red", linetype = "dashed")
```

```{r}
dwebble = dt[dt$pokemon == "Crustle",c('level', 'score', 'interrupt','damage_done','damage_healed', 'win')]
glm.fit <- glm(win ~ level + score + damage_healed, data = dwebble, family=binomial(link='logit'))
summary(glm.fit)
```

```{r}
MrMime = dt[dt$pokemon == "MrMime",c('level', 'score', 'interrupt','damage_done','damage_healed', 'win')]
glm.fit <- glm(win ~ level + score  + damage_healed, data = MrMime, family=binomial(link='logit'))
summary(glm.fit)
```

```{r}

library(leaps)
regfit.fwd=regsubsets(win ~ level + kill + assist + interrupt + damage_done + damage_taken + damage_healed, data=dt, method="forward")

summary(regfit.fwd)
```
